### 2016-02-20 doc/manuscript.tex
clean up introduction with RK changes

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[a4paper]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Extracellular potentials generated by axonal projections are shaped by patterns of bifurcations and terminations},
            pdfauthor={Thomas McColgan, Hermann Wagner, Catherine E Carr, Richard Kempter},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\usepackage{setspace}
\doublespacing

\title{Extracellular potentials generated by axonal projections are shaped by
patterns of bifurcations and terminations}
\author{Thomas McColgan, Hermann Wagner, Catherine E Carr, Richard Kempter}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}\label{introduction}

Extracellular field potentials (EFPs) are at the heart of many
experimental methods used to examine the inner workings of the brain.
This includes invasive methods such as Local Field Potentials (LFP),
Current Source Density (CSD) and Multiunit Activity (MUA) as well as
noninvasive methods such as Electroencephalogram (EEG),
Electrocorticogram (ECoG) and the auditory brainstem response (ABR)
(Nunez and Srinivasan, 2006; Brette and Destexhe, 2012). The origins of
these EFPs, especially in cases in which the activity is not clearly
attributable to a single cell, is a matter of debate.

EFPs in the brain were long thought to be primarily synaptic in origin
(Buzsáki et al., 2012). As a consequence, many modeling studies focussed
on the extracellular fields induced by synaptic currents on the
dendrites and the soma of a postsynaptic neuron (Holt and Koch, 1999;
Gold et al., 2006; Lindén et al., 2010, 2011; Einevoll et al., 2013;
Fernández-Ruiz et al., 2013). However, a number of recent data analysis
and modeling efforts have revealed that active, non-synaptic membrane
currents can play an important role in generating EFPs (Ray and
Maunsell, 2011; Belluscio et al., 2012; Schomburg et al., 2012; Reimann
et al., 2013; Anastassiou et al., 2015), including far reaching
potentials detectable at the scalp (Teleńczuk et al., 2011, 2015).
However, currents from the axon are still thought to be so small as to
be of minor importance for the EFP.

One of the reasons for the traditional assumption that axonal currents
contribute little to the EFP is that the far field of an action
potential traveling along an idealized straight axon is quadrupolar,
meaning that its far field decays faster with distance than synaptic
sources, which are thought to be dipolar (Nunez and Srinivasan, 2006).
However, in experimental studies that recorded the EFP of axonal
responses, the EFP often has a dipolar structure. For example, Blot and
Barbour (2014) report an EFP with the characteristic dipolar structure
recorded from the vicinity of cerebellar Purkinje cell axons.

\begin{itemize}
\tightlist
\item
  find more examples in literature
\end{itemize}

In the results section we will present a further new example of such a
field potential from the auditory brainstem of the barn owl.

This discrepancy between theory and experiment raises the open question
of how axons are able to generate dipolar field potentials.

The aim of this study is to show how dipolar far fields in the EFP of
axons can be explained by the axons' anatomical structure. In
particular, we show that the branchings and termonations of axons in
their projection area lead to a dipolar EFP structure. Such axon
bundles, often also called nerves or fascicles, exist throughout the
peripheral and central nervous system of most animals (Nornes and Das,
1972; Goodman et al., 1984; Hentschel and Ooyen, 1999; Kandel et al.,
2000). The white matter of the mammalian brain can be viewed as an
agglomeration of such bundles (Schüz and Braitenberg, 2002). We
therefore predict pronounced contributions of nerves or fascicles to
EFPs, which are neglected in all current models.

Properties of the EFP due to axonal projections are first investigated
by a numerical model based on forward simulation (Holt and Koch, 1999;
Gold et al., 2006). This first model includes a large-scale
multicompartment model of the axon population. We then show a concrete
example of the effects in experimental data using as a set of multi-site
in-vivo electrophysiological recordings from the barn owl brain stem.
Finally, we outline the mechanisms underlying the observed effects by
means of a second, analytically tractable, model of a generic axon
bundle.

\section{Results}\label{results}

\emph{Axonal projections generate a dipole-like field potential.} To
understand how axons can generate a dipolar EFP, we simulated a parallel
fiber bundle that at first runs at constant diameter without
bifurcations and then reaches a projection zone. Within this projection
zone, the fibers first bifurcate, which increases the number of fibers.
Finally, the axons reach a termination zone, in which the number of
fibers decreases to zero (Figure 1A).

The spiking activity of a generic axon bundle was simulated by
stimulating the fibers with a background spontaneous firing rate of 100
spikes/second, upon which a pulse of increased activity was added. This
pulse of activity could relate to any kind of evoked activity in the
nervous system, such as a sensory stimulation, motor activity or a
spontaneous transient increase in population spiking activity. We chose
a gaussian pulse with an amplitude of 2900 spikes/second and a standard
deviation of 2.8 ms.

To characterize the spatiotemporal dynamics of the evoked EFP, the
potential was calculated for several locations along the axon trunk.
Because we first want to study the effect of the rate pulse on the EFP,
we low-pass filtered the EFP responses with a cut-off frequency of 1 kHz
(Figure 1B). The distribution of maximum amplitudes of these responses
are shown by the colored contour lines in Figure 1A and the colored
traces in Figure 1B. Surrounding the projection zone of the axon bundle,
amplitudes show the double-lobed shape typical of a dipole.

In Figure 1B, all the responses showed a biphasic deflection that was
elicited by the population firing rate pulse. This deflection was
visible at all recording locations. The biphasic deflection reverses
polarity in the middle of the bifurcation zone of the bundle (red box in
Figure 1B). The polarity reversal occurs by a decrease of the amplitude
to zero and subsequent reappearance with reversed polarity, as opposed
to a polarity reversal through a gradual shift in phase. This behaviour
is also typical for a dipolar field potential.

The point of the polarity reversal coincides roughly with the middle of
the projection zone. This means that the absolute value of the response
amplitude reaches a local minimum just when the number of axonal fibers
reaches a maximum. To better understand how anatomical features and
response amplitude are related, we compared the signed maximum value
(meaning the \emph{signed} value corresponding to the maximum
\emph{absolute} amplitude) of the EFP with the difference in the density
of branchings and terminations (Figure 1C). Interestingly, the amplitude
of the maximum response (red trace) follows the distribution of
branchings and terminations (black histogram).

Along the nerve trunk the fiber density is constant, and the amplitude
is unchanged. As the axon bundle reaches its projection zone, the number
of bifurcations increases, and the response increases in amplitude. In
the middle of the projection zone, the number of bifurcations and
terminations cancels out. At the same depth, the amplitude crosses zero.
As the end of the projection zone approaches, the terminations outweigh
the bifurcations, and the amplitude increases, but the polarity is
reversed. This transition also corresponds to the traces in the red box
in Figure 1B. As the axon bundle ends, there are no longer any
bifurcations or terminations, and the amplitude decays toward zero.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_1.pdf}
\caption{Axonal projections generate a dipole-like extracellular field
potential (EFP) when stimulated with a pulse of activity. \textbf{A}
shows the modeled axon bundle in black, along with iso-amplitude lines
for the low-pass filtered EFP signature of the activity pulse. The
contours show the typical double-lobe of a dipole. The waveforms,
recorded at the locations of the colored dots in \textbf{A} and shown in
\textbf{B}, also show a polarity reversal. The most salient locations
are highlighted with the red box. The transition ocurrs by inverting the
envelope with aproximately constant phase, not by a shifting of phase
with constant envelope. \textbf{C} shows the progression of the maximum
amplitude with depth (red line). It closely follows the local density of
bifurcations and terminations (black histogram). The situation is
different for the high-frequency component. The iso-amlitude contours
shown in \textbf{D} do not exhibit the same double-lobe structure, and
the waveforms (\textbf{C}) also don't reverse polarity. The amplitude
envelope of the high-frequency component (red line in \textbf{E})
follows the density of fibers (black histogram in \textbf{E}).}
\end{figure}

To understand the contributions related to individual spikes, we next
turned our attention to the high-frequency range. We filtered the EFP
responses with a high-pass filter, using the same cut-off frequency of
1kHz as before. The maximum amplitudes for this frequency band did not
show a double-lobe, but rather appeared like an ellipsoid centered at
the projection zone (Figure 1D). The waveforms are more disparate across
recording locations, suggesting a more local nature. There is no visible
polarity reversal (Figure 1E).

The high-frequency component changes its amplitude in the longitudinal
direction in accordance to the local fiber density (Figure 1F). It has
an approximately constant amplitude along the nerve trunk and then
increases in amplitude as the fiber density is increased by
bifurcations. As the fibers terminate and fiber density decreases, so
does the amplitude of the high-frequency component.

We have thus shown a qualitatively different behaviour in the low- and
high-frequency components of the EFP. Both appear to be related to
anatomy, though each in a different way. Together, these relationships
suggest that the particular branching and terminating structure of the
axon bundle gives rise to the dipolar field potential. Because dipolar
fields are associated with a longer spatial reach than the quadrupolar
fields typically associated with axonal spikes, we decided to
investigate the effect that axon structure has on the spatial reach of
the frequency components (Figure 2). To do this we performed a further
simulation of an axon bundle, this time omitting the bifurcations, and
letting the fibers terminate directly. In axial direction, the
low-frequency component has a further reach than the high frequency
component in the bifurcating case (Figure 2A), but not in the
non-bifurcating case (Figure 2C). In the radial direction, the situation
is different: the bifurcating (Figure 2B) and non-bifurcating (Figure
2D) case both show a faster decay of the high-frequency component than
of the low-frequency component. To assure that the effect was due to the
bifurcations themselves and not due to the fanned out structure of the
bifurcating bundle seen in Figures 2A and 2D, we here changed the layout
of both bundles to have constant widths.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_2.pdf}
\caption{Low-frequency component of the axon bundle EFP exceeds
high-frequency component in reach, modulated by axon structure. All
plots show the scaling of the low-frequency (blue lines) and
high-frequency (green lines) frequency components on a
doubly-logarithmic scale. Amplitudes were normalized to unit value at
the end of the tuft in axial direction (\textbf{A} and \textbf{B}) and
from the trunk in radial direction (\textbf{C} and \textbf{D}).
\textbf{A} and \textbf{C} show the behaviour in the case of a
bifurcating bundle, while \textbf{B} and \textbf{D} show the
nonbifurcating case. In the bifurcating case, the low-frequency
component exceeds the reach of the high-frequency comonent in both
directions. In the non-bifurcating case, the decay is the same for both
components in the axial direction. This is consistent with the presence
of a low-frequency dipole-like potential in the bifurcating bundle but
not in the non-bifurcating bundle.}
\end{figure}

\emph{The barn owl neurophonic as an example that shows these
properties.} In order to test our model of the extracellular field of
axon bundles, we compared it to recordings from the barn owl auditory
brain stem. We adjusted the model parameters to match the anatomy and
physiology of nucleus laminaris (NL), and set the activation statistics
to match measured PSTHs in response to stimulation by a click sound
presented to one ear. We then simulated the resulting EFP at varying
depths, corresponding to recording depths in the experiment. The
resulting recordings from experiment and simulation are shown in Figure
3.

The simulated EFPs showed several characteristics that are also observed
in the data. Firstly, the high-frequency component shows a steady
increase in latency along the projections' depth, and has its maximum in
amplitude in the middle of NL. The low-frequency component reverses
polarity along the depth of NL, and almost vanishes in the middle of NL.
This is the same behaviour as shown for the generic axon bundle in the
previous section.

It is interesting to note that a similar reversal of polarity in NL has
been reported for chicken (Schwarz, 1992), as well as in the mammalian
analog to NL, the medial superior olive (Laughlin et al., 2010). The
phase reversal in this case has been modeled based on the assumption
that the postsynaptic NL and MSO dendrites with their bipolar morphology
have a dipolar EFP response (Laughlin et al., 2010; Goldwyn et al.,
2014). However, in the owl this dipolar morphology is largely absent
(Carr and Konishi, 1990), making dendritic sources unlikely.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/fig_3.pdf}
\caption{Data from the barn owl shows the expected behaviour predicted
by the model. (\textbf{A},\textbf{B}) shows data from the barn owls
nucleus laminaris in response to an auditory click stimulus, compared to
a simulation of the axonal structure and activation in
(\textbf{C},\textbf{D}). The click stimulus induces a pulse of activity
in the afferent axon bundle. The low-frequency components (\textbf{A}
and \textbf{C}) show the polarity reversal. The high frequency component
(\textbf{B} and \textbf{D}), does not show such a reversal, but rather
shows a steady increase in phase with depth.}
\end{figure}

\emph{Mechanism underlying the observed properties I.} We begin by
looking at the response of a single axon tree to a single spike at
various locations along the tree, as shown in Figure 4. This can already
give us a rough idea about the origin of the observed polarity reversal.

The extracellular response at a recording location along a continuation
of the axon, i.e.~not close to a bifurcation or termination, is
triphasic (blue trace in Figure 4B). It has been noted that this shape
is approximated by the second derivative of the membrane voltage: the
first positive peak is the initial bend to the rising flank of the
action potential, the main dip corresponds to the high curvature at the
peak of the action potential, and the final, positive peak corresponds
to the bend from hyperpolarization to recovery.

The potential near a bifurcation (red trace in Figure 4B) has a more
biphasic shape. There is a small initial peak, but the response is
dominated by the second, negative and the third, positive peak. This can
be understood by comparing the shape to the previous example along a
non-bifurcating axon: The initial shape resembles the non-bifurcating
case, because the action potential is still mainly within the segments
before the bifurcation. As the action potential passes the bifurcation,
there are now two instances of the action potential, and the second and
third peak are doubled in size, leading to the more biphasic response.

The same reasoning can be applied to the case of the biphasic response
at a termination (green trace in Figure 4B): here there is an
approaching action potential, leading to an initial positive peak, but
as the termination is reached, the current flow ceases, and there is
only a reduced second peak, and the third peak, associated with the
action potential moving away from the recording location, disappears
completely.

A notable point here is that the potential at the bifurcation and
termination are similar in shape, but opposite in polarity. This
resembles the observed polarity reversal. It is thus not hard to imagine
that this effect is caused by the response being dominated by
bifurcations in the early part of NL where the axons branch out, and
then by terminations in the later part, where terminations are more
prevalent, with the two canceling out in the middle. We will test this
intuition in the following by examining the properties of a full bundle
of axons in a simplified analytical model.

\begin{figure}[htbp]
\centering
\includegraphics{../figs/mockups/fig4.pdf}
\caption{The dipolar behaviour can be understood by examining individual
action potentials on a single axon tree. Comparing the low frequency owl
data (\textbf{A}) to a single axon and action potential in model
(\textbf{B}) shows a similar behaviour. In particular, the potential at
a termination and that at a bifurcation (red and green curves in
\textbf{B}) are approximately inverted.}
\end{figure}

\emph{Mechanism underlying the observed properties II.} In the case of a
simplified one-dimensional axon bundle, the EFP at a given location may
be described by the formula
\(\Phi(\mathbf{r},t) = \left[\left\{n(z)\cdot i_\lambda(z,t)\right\}\ast w(a,z)\right]_z\)
(see Methods and appendix). Assuming that the weighting function \(w\)
tends to 0 for large \(\left|z\right|\), and with the introduction of
the single axon current integral
\(I_\lambda(z,t)=\int_{-\infty}^zi_\infty(z',t)dz'\), the equation can
be split (for a derivation, see appendix) into two components:

\begin{align}
  \label{eqn:splitpot} 
  \Phi(\mathbf{r},t) = &\left[\left\{n(z)\cdot I_\lambda(z,t)\right\}\ast w'(\rho,z)\right]_z \\
  & - \left[\left\{n'(z)\cdot I_\lambda(z,t)\right\}\ast w(\rho,z)\right]_z 
\end{align}

In the first component the single axon current integral
\(I_\lambda(z,t)\) is multiplied by the local density of the fibers
\(n\), while the weighting function is replaced by its spatial
derivative \(w'\). In the second component, \(I_\lambda\) is multiplied
by the derivative of the density \(n'\), but here the weighting function
\(w\) is unchanged. The functions \(w\) and \(w'\) are shown in the time
and frequency domain in Figure 4.

Further inspection of the two components shows an intriguing similarity
to the properties of the response observed in the previous sections. The
first component is high-pass filtered, and is related to the fiber
density, just like `noise' component in Figure 1 and the high-frequency
component in the barn owl data and model. The second component is
low-pass filtered, and related to the derivative of the fiber density,
just like the firing rate pulse related deflection in Figure 1, and the
low-frequency component in the barn owl.

\begin{itemize}
\tightlist
\item
  Add simplified explanations of:
\item
  spatial filtering
\item
  dipole/stationary potentials
\end{itemize}

\begin{figure}[htbp]
\centering
\includegraphics{../figs/mockups/fig5.pdf}
\caption{Analytical model of the axon bundle potential explains the
effects observed in the numerical model and example data. (\textbf{A})
shows the behaviour of a simplified fiber bundle with a piecewise
constant fiber density (\textbf{Aa}). The high-frequency component
(\textbf{Ab}) shows no polarity reversal, while the low-frequency
component (\textbf{Ac}) does, as expected from the data and modeling.
This can be understood by decomposing the signal into two components.
The first component is governed by the bifurcation and termination
density, and is filtered by the regular weighting function
(\textbf{Ba}), which acts as a low-pass filter (\textbf{Bb}). The second
component is governed by the fiber density, and is filtered by the
derivative of the weighting function (\textbf{Bc}), which acts as a
high- or band-pass filter (\textbf{Bd}).}
\end{figure}

\section{Discussion}\label{discussion}

\begin{itemize}
\tightlist
\item
  Relevance of Findings

  \begin{itemize}
  \tightlist
  \item
    Interpretation of CSD

    \begin{itemize}
    \tightlist
    \item
      Classical CSD: constant fiber density, variable currents
    \item
      Here: variable fiber density, constant currents
    \end{itemize}
  \item
    Dipole has far field, ABR response?
  \end{itemize}
\item
  Compare to other auditory systems (Chicken NL, MSO)

  \begin{itemize}
  \tightlist
  \item
    Speculate on functional relevance of polarity shift (a la Rinzel \&
    Goldwyn, Weiss 2010, Einevoll\&Panzeri review)
  \end{itemize}
\item
  Two simultaneous modes: stationary dipole and continuous (cf Mc
  Laughlin, etc).
\item
  compare to other fiber bundle systems
\end{itemize}

\section{Methods}\label{methods}

\subsection{Experimental recordings}\label{experimental-recordings}

\subsection{Numerical model}\label{numerical-model}

We modeled the axons using NEURON (Hines and Carnevale, 1997; Hines et
al., 2009) in a model based on previous work by (\textbf{refs}),
including the high and low threshold potassium channels used by
(\textbf{refs}). The axon was modeled as a sequence of active nodes and
passive myelinated segments. The parameters for the active and passive
segments are described in Table \ref{tab:modparam}. Unlike previous
models, we included branching axons in our simulations. These were
generated by connecting two passive segments to a node, and continuing
the alternation of active and passive segment in each resulting branch.

Action potentials were initiated in the axons by injecting a current
pulse into the first node of Ranvier of the axon. The times of the
injections were chosen by drawing from an inhomogeneous Poisson
distribution based on the firing statistics of NM units in response to a
given stimulus.

\begin{itemize}
\tightlist
\item
  Split barn owl and general axon bundle
\end{itemize}

Axon branching patterns were generated procedurally, \textbf{better
wording} starting with the root segment, placed at the border of NL. In
order to avoid artifacts from the current pulse injection and to
simulate the fiber tract leading up to NL, a sequence of 10 active and
passive segments without bifurcations was added before the root. To this
root, segments were appended iteratively. Before adding a segment, a
decision whether to branch or terminate an axon was drawn from a
probability distribution dependent on the dorsoventral depth of the end
of the previous segment. These probability distributions were modeled as
logistic functions with the parameters adjusted to roughly match the
numbers of branchings and terminations found in the tracings shown in
publications such as Carr and Konishi (1990). This meant that an initial
phase of bifurcations was followed by a phase of terminations, with the
probability of termination reaching 100\% at the end of NL.

The simulations of these axons yielded the membrane currents, from which
we calculated the extracellular fields. The procedure for this is
described in detail by Gold et al. (2006), among others. Briefly, the
extracellular medium is assumed to be a homogeneous volume conductor
with conductivity \(\sigma_e\), and a quasi-static approximation of the
electrical field potential \(\phi\) is made. The extracellular potential
due to a current distribution \(i(\mathbf{r},t)\) is then governed by
the equation \(\Delta \phi = \frac{1}{\sigma_e} i(\mathbf{r},t)\), with
\(\Delta\) denoting the Laplace operator. If the currents \(i\) are
constrained to a volume \(V\), this equation has the solution:

\begin{equation}
  \phi(\mathbf{r},t)=\frac{1}{4\pi\sigma_{e}}\int_{V}\frac{i(\mathbf{r}',t)}{|\mathbf{r}-\mathbf{r}'|}\textrm{d}\mathbf{r}'
  \label{eqn:basic}
\end{equation}

Since the currents though the myelinated segments were negligible, and
the nodes of Ranvier are small, we used the point source approximation
of only the node currents, and did not include the line source
approximation used by Holt and Koch (1999).

\begin{table}[h]
  \begin{centering}
    \begin{tabular}{|c|c|}
      \hline 
      parameter & value\tabularnewline
      \hline 
      \hline 
      $g_{\text{Na}}$ & 0.8 S/cm\texttwosuperior{}\tabularnewline
      \hline 
      $g_{\text{KLVA}}$ & 0.1 S/cm\texttwosuperior{}\tabularnewline
      \hline 
      $g_{\text{KHVA}}$ & 1.5 S/cm\texttwosuperior{}\tabularnewline
      \hline 
      $C_{\text{mem}}$ & 1.0 $\mu$F/cm\texttwosuperior{}\tabularnewline
      \hline 
    \end{tabular}
    \par
  \end{centering}
  \caption{Model Parameters **add: leak conductance, myelin parameters, extracellular conductivity sizes**}
  \label{tab:modparam}
\end{table}

\subsection{Simplified axon bundle
model}\label{simplified-axon-bundle-model}

\label{sec:efpresp}

We define the spatial dimension in cylindrical coordinates
\(\mathbf{r}=(\rho,\varphi,z)\) such that \(z\) is the dorsoventral
direction, increasing from dorsal to ventral, and \(z=0\) is the dorsal
border of NL. We further define \(\hat{\mathbf{e}}_{z}\) as the unit
vector in \(z\) direction, and \(\hat{\mathbf{e}}_{\rho}\) as an
arbitrary unit vector perpendicular to the \(z\) direction.

Let us first consider a simple model axon that extends infinitely on
both sides and in a straight line from the dorsal direction into NL, and
at 0 in the remaining coordinates. Furthermore, we consider a single
action potential propagating along this axon. This axon then has a
non-zero current density only at \(\rho=0\), which we denote
\(i_{\infty}(z,t)\), meaning that in this case
\(i(\mathbf{r},t)=\delta(\rho)i_{\infty}(z,t)\). Applying this to
equation \ref{eqn:basic}, corresponding response
\(\kappa_{\infty}(\mathbf{r},t)\) of an action potential propagating
through such a line-axon is then

\begin{align}
\kappa_{\infty}(\mathbf{r},t) & =\frac{1}{4\pi\sigma_{e}}\int_{V}\frac{\delta(\rho')i_{\infty}(z',t)}{|\mathbf{r}-\mathbf{r}'|}\textrm{d}\mathbf{r}'\label{eq:halfline}\\
& =\frac{1}{4\pi\sigma_{e}}\int_{-\infty}^{\infty}\frac{i_{\infty}(z',t)}{|\mathbf{r}-z'\hat{\mathbf{e}}_{z}|}\textrm{d}z'
\end{align}

We will call this response a spike kernel because it will take the role
of an integral kernel in the following.

Due to the rotational symmetry, the kernel at a distance \(\rho\) from
the axon, regardless of \(\varphi\), can then be described by

\begin{align} \label{eqn:infkern}
\kappa_{\infty}(\mathbf{r},t) & =\frac{1}{4\pi\sigma_{e}}\int_{-\infty}^{\infty}\frac{i_{\infty}(z',t)}{|(z-z')\hat{\mathbf{e}}_{z} + \rho\hat{\mathbf{e}}_{\rho}|}\textrm{d}z' \\
& =\frac{1}{4\pi\sigma_{e}}\int_{-\infty}^{\infty}\frac{i_{\infty}(z',t)}{\sqrt{(z-z')^2 + \rho^2}}\textrm{d}z' 
\end{align}

Equation \ref{eqn:infkern} has the form of a convolution with a
weighting function \(w\):

\begin{equation}
  w(\rho,z) = \frac{1}{4\pi\sigma_e}\frac{1}{\sqrt{z^2+\rho^2}}
  \label{eqn:weighting}
\end{equation}

The convolution can then be written as:

\begin{equation}
\kappa_{\infty}(\mathbf{r},t) = \left[i_{\infty}(z,t)\ast w(\rho,z)\right]_z
\end{equation}

with the operator \(\left[\cdot\ast\cdot\right]_z\) denoting the
convolution with respect to the variable \(z\).

The simplest model of a terminating axon will be a semi-infinite axon,
where the membrane current flow is simply set to zero beyond the
termination point \(z_\text{term}\). Using the Heaviside step function
\(H\) we find kernel \(\kappa_{\text{term}}\) of a spike in a
terminating axon:

\begin{equation}
  \kappa_{\text{term}}(\mathbf{r},t) = \left[\left\{H(z_\text{term}-z)\cdot i_{\infty}(z,t)\right\}\ast w(\rho,z)\right]_z
\end{equation}

Similarly, the model of a bifurcating axon would be one in which the
current flow after the bifurcation point \(z_\text{bif}\) is double that
of the single infinite fiber.

\begin{equation}
  \kappa_{\text{bif}}(\mathbf{r},t) = \left[\left\{\left(1+H(z-z_\text{bif})\right)\cdot i_{\infty}(z,t)\right\}\ast w(\rho,z)\right]_z
\end{equation}

This places the child branches in superposition at \(\rho=0\), which is
a useful approximation for small branching angles.

Approximating terminations and bifurcations in this way disregards the
boundary effects that may appear due to the conservation of charge and
the inability of charge to flow beyond the termination. However, since
the integral over \(i_\infty\) is zero for realistic action potentials,
the conservation of charge is maintained in the long run.

In more general terms, a coherently stimulated axon bundle with
terminations and bifurcations along it's path can be described by the
number of individual fibers \(n(z)\) at any depth. The extracellular
potential \(\kappa_\text{b}\) of such a bundle in which the action
potential is initiated at the same time and location is then given by

\begin{equation}
  \kappa_{\text{b}}(\mathbf{r},t) = \left[\left\{n(z)\cdot i_{\infty}(z,t)\right\}\ast w(\rho,z)\right]_z
  \label{eqn:bundlekern}
\end{equation}

If we want to calculate the combined field of many axons and action
potentials, we can do so by linearly summing the individual axon fields.

If all axons in the bundle are driven with the same the mean firing rate
\(\lambda(t)\), which might be modulated in time, the field potential
\(\Phi(\mathbf{r},t)\) in response to this firing rate will be the
(temporal) convolution of the firing rate with the response of the
bundle to a coherent spike in all axons:

\begin{equation}
  \Phi(\mathbf{r},t) = \left[\kappa_\text{b}(\mathbf{r},t) \ast \lambda(t)\right]_t
  \label{eqn:fullpot}
\end{equation}

Substituting equation \ref{eqn:bundlekern} into equation
\ref{eqn:fullpot}, and taking advantage of the fact that only
\(i_\infty\) and \(\lambda\) depend on \(t\) gives

\begin{align}
  \label{eqn:switchedpot}
  \Phi(\mathbf{r},t) & =  \left[\left[\left\{n(z)\cdot i_{\infty}(z,t)\right\}\ast w(\rho,z)\right]_z \ast \lambda(t)\right]_t \\
  & =  \left[\left\{n(z)\cdot \left[i_{\infty}(z,t)\ast \lambda(t)\right]_t\right\}\ast w(\rho,z)\right]_z
\end{align}

With the average current in a single infinite fiber stimulated with
\(\lambda\), which we will denote it with
\(i_\lambda(z,t) := \left[i_{\infty}(z,t)\ast \lambda(t)\right]_t\) this
gives us

\begin{align}
  \Phi(\mathbf{r},t) & =  \left[\left\{n(z)\cdot i_\lambda(z,t)\right\}\ast w(a,z)\right]_z
  \label{eqn:switchedpotend}
\end{align}

\section*{Bibliophraphy}\label{bibliophraphy}
\addcontentsline{toc}{section}{Bibliophraphy}

\hypertarget{refs}{}
\hypertarget{ref-Anastassiou2015Cell}{}
Anastassiou CA, Perin R, Buzsáki G, Markram H, Koch C (2015) Cell type-
and activity-dependent extracellular correlates of intracellular
spiking. Journal of Neurophysiology 114:608--623 Available at:
\url{http://dx.doi.org/10.1152/jn.00628.2014}.

\hypertarget{ref-Belluscio2012CrossFrequency}{}
Belluscio MA, Mizuseki K, Schmidt R, Kempter R, Buzsáki G (2012)
Cross-Frequency Phase--Phase Coupling between Theta and Gamma
Oscillations in the Hippocampus. The Journal of Neuroscience 32:423--435
Available at: \url{http://dx.doi.org/10.1523/jneurosci.4122-11.2012}.

\hypertarget{ref-Blot2014Ultrarapid}{}
Blot A, Barbour B (2014) Ultra-rapid axon-axon ephaptic inhibition of
cerebellar Purkinje cells by the pinceau. Nat Neurosci 17:289--295
Available at: \url{http://dx.doi.org/10.1038/nn.3624}.

\hypertarget{ref-Brette2012Handbook}{}
Brette R, Destexhe A eds. (2012) Handbook of Neural Activity
Measurement. Cambridge: Cambridge University Press. Available at:
\url{http://dx.doi.org/10.1017/cbo9780511979958}.

\hypertarget{ref-Buzsaki2012Origin}{}
Buzsáki G, Anastassiou CA, Koch C (2012) The origin of extracellular
fields and currents--EEG, ECoG, LFP and spikes. Nature reviews
Neuroscience 13:407--420 Available at:
\url{http://dx.doi.org/10.1038/nrn3241}.

\hypertarget{ref-carr90}{}
Carr CE, Konishi M (1990) A circuit for detection of interaural time
differences in the brain stem of the barn owl. The Journal of
Neuroscience 10:3227--3246 Available at:
\url{http://www.jneurosci.org/content/10/10/3227.abstract}.

\hypertarget{ref-Einevoll2013Modelling}{}
Einevoll GT, Kayser C, Logothetis NK, Panzeri S (2013) Modelling and
analysis of local field potentials for studying the function of cortical
circuits. Nature reviews Neuroscience 14:770--785 Available at:
\url{http://dx.doi.org/10.1038/nrn3599}.

\hypertarget{ref-FernandezRuiz2013Cytoarchitectonic}{}
Fernández-Ruiz A, Muñoz S, Sancho M, Makarova J, Makarov VA, Herreras O
(2013) Cytoarchitectonic and Dynamic Origins of Giant Positive Local
Field Potentials in the Dentate Gyrus. The Journal of Neuroscience
33:15518--15532 Available at:
\url{http://dx.doi.org/10.1523/jneurosci.0338-13.2013}.

\hypertarget{ref-Gold2006Origin}{}
Gold C, Henze DA, Koch C, Buzsáki G (2006) On the origin of the
extracellular action potential waveform: A modeling study. Journal of
neurophysiology 95:3113--3128 Available at:
\url{http://dx.doi.org/10.1152/jn.00979.2005}.

\hypertarget{ref-Goldwyn2014Model}{}
Goldwyn JH, Laughlin MM, Verschooten E, Joris PX, Rinzel J (2014) A
Model of the Medial Superior Olive Explains Spatiotemporal Features of
Local Field Potentials. The Journal of Neuroscience 34:11705--11722
Available at: \url{http://dx.doi.org/10.1523/jneurosci.0175-14.2014}.

\hypertarget{ref-Goodman1984Cell}{}
Goodman CS, Bastiani MJ, Doe CQ, Lac S du, Helfand SL, Kuwada JY, Thomas
JB (1984) Cell recognition during neuronal development. Science (New
York, NY) 225:1271--1279 Available at:
\url{http://view.ncbi.nlm.nih.gov/pubmed/6474176}.

\hypertarget{ref-Hentschel1999Models}{}
Hentschel HGE, Ooyen A van (1999) Models of axon guidance and bundling
during development. Proceedings of the Royal Society of London B:
Biological Sciences 266:2231--2238 Available at:
\url{http://dx.doi.org/10.1098/rspb.1999.0913}.

\hypertarget{ref-Hines1997NEURON}{}
Hines ML, Carnevale NT (1997) The NEURON Simulation Environment. Neural
Computation 9:1179--1209 Available at:
\url{http://dx.doi.org/10.1162/neco.1997.9.6.1179}.

\hypertarget{ref-Hines2009NEURON}{}
Hines ML, Davison AP, Muller E (2009) NEURON and Python. Frontiers in
neuroinformatics 3 Available at:
\url{http://dx.doi.org/10.3389/neuro.11.001.2009}.

\hypertarget{ref-Holt1999Electrical}{}
Holt GR, Koch C (1999) Electrical Interactions via the Extracellular
Potential Near Cell Bodies. Journal of Computational Neuroscience
6:169--184 Available at:
\url{http://dx.doi.org/10.1023/a:1008832702585}.

\hypertarget{ref-kandel2000principles}{}
Kandel ER, Schwartz JH, Jessell TM, Others (2000) Principles of neural
science. McGraw-Hill New York.

\hypertarget{ref-Laughlin2010Oscillatory}{}
Laughlin MM, Verschooten E, Joris PX (2010) Oscillatory Dipoles As a
Source of Phase Shifts in Field Potentials in the Mammalian Auditory
Brainstem. The Journal of Neuroscience 30:13472--13487 Available at:
\url{http://dx.doi.org/10.1523/jneurosci.0294-10.2010}.

\hypertarget{ref-Linden2010Intrinsic}{}
Lindén H, Pettersen K, Einevoll G (2010) Intrinsic dendritic filtering
gives low-pass power spectra of local field potentials. Journal of
computational neuroscience 29:423--444 Available at:
\url{http://dx.doi.org/10.1007/s10827-010-0245-4}.

\hypertarget{ref-Linden2011Modeling}{}
Lindén H, Tetzlaff T, Potjans TC, Pettersen KH, Grün S, Diesmann M,
Einevoll GT (2011) Modeling the Spatial Reach of the LFP. Neuron
72:859--872 Available at:
\url{http://dx.doi.org/10.1016/j.neuron.2011.11.006}.

\hypertarget{ref-Nornes1972Temporal}{}
Nornes HO, Das GD (1972) Temporal pattern of neurogenesis in spinal
cord: cytoarchitecture and directed growth of axons. Proceedings of the
National Academy of Sciences of the United States of America
69:1962--1966 Available at:
\url{http://view.ncbi.nlm.nih.gov/pubmed/4114859}.

\hypertarget{ref-Nunez2006Electric}{}
Nunez PL, Srinivasan R (2006) Electric Fields of the Brain. Oxford
University Press. Available at:
\url{http://dx.doi.org/10.1093/acprof:oso/9780195050387.001.0001}.

\hypertarget{ref-Ray2011Different}{}
Ray S, Maunsell JHR (2011) Different Origins of Gamma Rhythm and
High-Gamma Activity in Macaque Visual Cortex. PLoS Biol 9:e1000610+
Available at: \url{http://dx.doi.org/10.1371/journal.pbio.1000610}.

\hypertarget{ref-Reimann2013Biophysically}{}
Reimann MW, Anastassiou CA, Perin R, Hill SL, Markram H, Koch C (2013) A
Biophysically Detailed Model of Neocortical Local Field Potentials
Predicts the Critical Role of Active Membrane Currents. Neuron
79:375--390 Available at:
\url{http://dx.doi.org/10.1016/j.neuron.2013.05.023}.

\hypertarget{ref-Schomburg2012Spiking}{}
Schomburg EW, Anastassiou CA, Buzsáki G, Koch C (2012) The Spiking
Component of Oscillatory Extracellular Potentials in the Rat
Hippocampus. The Journal of Neuroscience 32:11798--11811 Available at:
\url{http://dx.doi.org/10.1523/jneurosci.0656-12.2012}.

\hypertarget{ref-Schuz2002Human}{}
Schüz A, Braitenberg V (2002) The Human Cortical White Matter. In:
Cortical areas (Schüz A, ed), pp 377--386. Abingdon, UK: Taylor \&
Francis. Available at:
\url{http://dx.doi.org/10.4324/9780203219911/_chapter/_16}.

\hypertarget{ref-Schwarz1992Can}{}
Schwarz DW (1992) Can central neurons reproduce sound waveforms? An
analysis of the neurophonic potential in the laminar nucleus of the
chicken. The Journal of otolaryngology 21:30--38 Available at:
\url{http://view.ncbi.nlm.nih.gov/pubmed/1564747}.

\hypertarget{ref-Telenczuk2011Highfrequency}{}
Teleńczuk B, Baker SN, Herz AVM, Curio G (2011) High-frequency EEG
covaries with spike burst patterns detected in cortical neurons. Journal
of Neurophysiology 105:2951--2959 Available at:
\url{http://dx.doi.org/10.1152/jn.00327.2010}.

\hypertarget{ref-Telenczuk2015Correlates}{}
Teleńczuk B, Baker SN, Kempter R, Curio G (2015) Correlates of a single
cortical action potential in the epidural EEG. NeuroImage 109:357--367
Available at: \url{http://dx.doi.org/10.1016/j.neuroimage.2014.12.057}.

\end{document}